Forecasting the 2024 U.S. Presidential Election Using a Model-Based Approach

### 2024-11-01 paper/paper.qmd
update paper

---
title: "Based on polling as of 27 October 2024, Harris expects to win on 49%, leads Trump by about 2% in 2024 US election"
subtitle: "Weighted Linear Regression Analysis of Aggregated Poll Data: Utilizing National Polls indicator, Population, and Pollster as Predictors."
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This paper predicts the outcome of the 2024 U.S. presidential election using a statistical model based on aggregated polling data for Kamala Harris and Donald Trump. We employ multi-level regression with post-stratification (MRP) using demographic predictors to estimate voter support. The analysis addresses polling biases and proposes an idealized survey methodology with a $100,000 budget. Our results offer insights into voter behavior and suggest improvements for future election forecasting models."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

```{r}
#| include: false
#| warning: false
#| message: false
harris_data <- read_parquet("~/2024 US Election Forecast/data/02-analysis_data/analysis_data_Harris.parquet")
trump_data <- read_parquet("~/2024 US Election Forecast/data/02-analysis_data/analysis_data_Trump.parquet")
```

@@

The structure of this paper is organized as follows: @sec-data provides details on the data sources and variables used in our analysis. @sec-model explains the modeling approach, including the assumptions and specifications of our linear regression framework. In @sec-results, we present our findings, emphasizing the key predictors of Harris's support. Finally, @sec-discussion explores the implications of our results and suggests potential directions for future research.
# Estimand

@@

In this study, we utilized several R packages to enhance our data manipulation, modeling, and visualization capabilities. The tidyverse package offered a comprehensive set of tools for data wrangling and analysis, improving workflow efficiency (@thereferencecanbewhatever). The here package aided in managing file paths, allowing for easy access to our data files (@citehere). We relied on janitor to perform data cleaning, as it provides functionalities to identify and rectify quality issues within the dataset (@citejanitor). For handling date-related operations, the lubridate package proved invaluable, simplifying the manipulation of time variables (@citelubridate). Lastly, arrow supported efficient data input and output in a performance-oriented format, essential for managing larger datasets (@citearrow). Our coding practices and file organization were informed by the structure outlined in @tellingstories.

@@

## Model

@@

```{r}
#| include: false
#| warning: false
#| message: false
harris_unweighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_harris_unweighted.rds")
harris_weighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_harris_weighted.rds")
trump_unweighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_trump_unweighted.rds")
trump_weighted_model <-readRDS(file = "~/2024 US Election Forecast/models/model_trump_weighted.rds")

@@

## Recent Three Months Support Trends Prediction

The recent three-month support trends for Kamala Harris and Donald Trump across key battleground states—Arizona, Georgia, Michigan, Nevada, North Carolina, Pennsylvania, and Wisconsin—show a highly competitive race. Each state’s panel displays support percentages over time from August to November 2024, with Harris shown in blue and Trump in red. The plot uses point size to represent poll weights, accounting for factors such as recency, pollster quality, and sample size, where larger points indicate higher-weighted polls, emphasizing more reliable data.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-support-trends-by-state
#| fig-cap: Support Trends for Kamala Harris and Donald Trump Across Selected States

harris_data <- harris_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

# Calculate weights for Trump data
trump_data <- trump_data %>%
  mutate(
    sample_size_weight = pmin(sample_size / 2300, 1),
    pollster_quality_weight = numeric_grade / 4,
    recency_weight = exp(-recency * 0.1)
  ) %>%
  group_by(pollster) %>%
  mutate(
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) %>%
  ungroup() %>%
  mutate(combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight)

harris_data <- harris_data %>% mutate(candidate = "Harris")
trump_data <- trump_data %>% mutate(candidate = "Trump")
combined_data_total <- bind_rows(harris_data, trump_data)

selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
recent_date_cutoff <- as.Date("2024-08-01")  # Set cutoff for recent data

harris_filtered <- harris_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)

trump_filtered <- trump_data %>%
  filter(state %in% selected_states, as.Date(end_date) >= recent_date_cutoff)

# Combine datasets for plotting with a new column for candidate
harris_filtered <- harris_filtered %>% mutate(candidate = "Harris")
trump_filtered <- trump_filtered %>% mutate(candidate = "Trump")
combined_data <- bind_rows(harris_filtered, trump_filtered)
# Plot with recent data, specific axis formatting, and smoother trend lines
ggplot(combined_data, aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  facet_wrap(~ state,ncol = 2,scales = "free_y") +
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Selected States",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-25"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate"))
```
@fig-support-trends-by-state illustrates the recent support trends for Kamala Harris and Donald Trump across selected battleground states from August to November 2024. Each subplot represents a different state, with support percentages for each candidate displayed over time. Key features include Harris shown in blue and Trump in red, with smoothed trend lines to capture changes over time. The plot's y-axis spans from approximately 30% to 50% support, while the x-axis highlights monthly intervals from August to November. For example, in Arizona, Trump maintains a slight lead with support fluctuating around 50% in October, compared to Harris around 48%. In Georgia, Trump's support reached just above 49% in late September, whereas Harris hovers closer to 47%. In Wisconsin, both candidates show tightly aligned trends near 48%, highlighting a close race. Each point's size reflects the poll weight—calculated using recency, pollster quality, and sample size—indicating the relative importance of each poll. This visualization emphasizes recent trends in candidate support across key battleground states.

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-national-support-trends
#| fig-cap: Support Trends for Kamala Harris and Donald Trump Across Selected States
# Combine datasets for plotting with a new column for candidate
harris_filtered <- harris_filtered %>% mutate(candidate = "Harris")
trump_filtered <- trump_filtered %>% mutate(candidate = "Trump")
combined_data <- bind_rows(harris_filtered, trump_filtered)

combined_data_total|>filter(national_poll==1)|>ggplot(aes(x = as.Date(end_date), y = pct, color = candidate, size = combined_weight)) +
  geom_point(alpha = 0.7) + # Adds points with transparency for overlapping data
  geom_smooth(aes(group = candidate), method = "loess", span = 0.2, se = FALSE, size = 1) + # Smoothed trend line
  labs(title = "Recent Support Trends for Kamala Harris and Donald Trump Across Country",
       x = "Date", y = "Support (%)") +
  scale_x_date(date_labels = "%b", date_breaks = "1 month", limits = as.Date(c("2024-08-01", "2024-10-27"))) +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_size_continuous(range = c(0.5, 3), guide = "none") +
  theme_minimal() +
  theme(
    legend.position = "top",
    axis.text.x = element_text(angle = 0, hjust = 0.5),
    panel.grid.major = element_line(size = 0.1, color = "grey80")
  ) +
  guides(color = guide_legend(title = "Candidate")) +  # Vertical line at October 25
  annotate("text", x = as.Date("2024-10-19"), y = Inf, label = "October 25th", vjust = 1, color = "black", size = 3.5) +
  geom_vline(xintercept = as.Date("2024-10-25"), linetype = "solid", color = "black", size = 0.8)   # Vertical line at October 25

```
@fig-national-support-trends illustrates the polling support trends for Kamala Harris and Donald Trump across the country from August to October. The scatter points represent individual poll results, with blue for Harris and red for Trump, while each dot's size reflects the weight of the poll, accounting for factors like sample size, quality, and recency. Harris’s support trend line generally remains above Trump’s, with an average support level around 49%, compared to Trump’s approximate average of 47% over this period. Toward the end of October, both candidates show a slight increase, with Harris’s support reaching just over 50% and Trump’s approaching 49%. This trend suggests a potential narrowing in support levels as the election approaches, providing a quantitative snapshot of national polling dynamics.

## Recent Three Months Support Trends Linear Model Prediction

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
set.seed(233)
new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 1000 
  ),
  national_poll = rep(c(0, 1), length(1000)),
  pollster = factor(sample(levels(harris_data$pollster), size = 1000, replace = TRUE), 
                    levels = levels(harris_data$pollster)),  
  population = factor(sample(c("lv","rv"), size = 1000, replace = TRUE)),
  sample_size = sample(1:2300, size = 1000, replace = TRUE),
  pollster_quality= sample(2:4, size = 1000, replace = TRUE)
)|>
  group_by(pollster)|>
  mutate(recency = as.numeric(difftime( as.Date("2024-11-05"),as.Date(end_date), units = "days")),
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>ungroup()|>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1),  # Set sample size; adjust accordingly

    # Pollster quality weight (assuming you want to use a dummy numeric grade, e.g., 2)
    pollster_quality_weight = pollster_quality / 4,

    recency_weight = exp(-recency * 0.1),  # Adjust decay rate as needed
    # Combine all weights into a final combined weight
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    candidate="Harris"
  )
#######
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 1000
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 1000, replace = TRUE), 
                    levels = levels(trump_data$pollster)),  
  population = factor(sample(c("lv","rv"), size = 1000, replace = TRUE)),
   national_poll = rep(c(0, 1), length(1000)),  
  sample_size =sample(1:2300, size = 1000, replace = TRUE),
  pollster_quality= sample(2:4, size = 1000, replace = TRUE)
)|>
  group_by(pollster)|>
  mutate(recency = as.numeric(difftime( as.Date("2024-11-05"),as.Date(end_date), units = "days")),
    recent_poll_count = sum(recency < 30),
    poll_frequency_weight = ifelse(recent_poll_count > 4, 4 / recent_poll_count, 1)
  ) |>
  ungroup() |>
  mutate(
    # Sample size weight (assuming a maximum of 2300 responses)
    sample_size_weight = pmin(sample_size / 2300, 1), 
    pollster_quality_weight = pollster_quality / 4,  
    recency_weight = exp(-recency * 0.1),  
    combined_weight = recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight,
    candidate="Trump"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| label: fig-support-predicted
#| fig-cap: Predicted Support Trends for Kamala Harris and Donald Trump using Fitted Linear Model
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

@@

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions<- bind_rows(new_data_harris, new_data_trump)
combined_predictions_filtered<-combined_predictions|>filter(end_date>=recent_date_cutoff)
# Determine the maximum end date for setting the x-axis limit
max_end_date <- as.Date("2024-10-30")

# Plotting with the filtered data
ggplot(combined_predictions_filtered, aes(x = end_date, color = candidate)) +
  geom_smooth(aes(y = predicted_unweighted, linetype = "Unweighted"), 
              method = "loess", se = TRUE, size = 1) +  # Unweighted model with smoothing
  geom_smooth(aes(y = predicted_weighted, linetype = "Weighted"), 
              method = "loess", se = TRUE, size = 1) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Weights: recency_weight * sample_size_weight * poll_frequency_weight * pollster_quality_weight\nLinear Regression Model: candidate support percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  scale_linetype_manual(name = "Linear Model Type", values = c("Unweighted" = "dashed", "Weighted" = "solid")) +
  theme_minimal() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1)) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))  +
  geom_vline(xintercept = as.Date("2024-10-25"), linetype = "solid", color = "black", size = 0.8) +  # Vertical line at October 25
  annotate("text", x = as.Date("2024-10-19"), y = Inf, label = "October 25th", vjust = 1, color = "black", size = 3.5)+  # Vertical line at October 25
  geom_text(aes(x = as.Date("2024-10-25"), y = 50, label = "49%"), color = "blue", hjust = -0.1, size = 3) +  # Harris label
  geom_text(aes(x = as.Date("2024-10-25"), y = 48, label = "48%"),color = "red", hjust = -0.1, size = 3)+  # Harris label
  geom_text(aes(x = as.Date("2024-10-25"), y = 46.5, label = "46.5%"), color = "red",hjust = -0.1, size = 3)  # Trump label
```

@@

# Additional data details {#first-model-details}

@@

# Model details {#sec-model-details}

@@

Harris model: RMSE of 3.44, indicating an average prediction error of
around 3.44 percentage points.\
Trump model: RMSE of 4.40, showing a slightly higher prediction error on
average.\

R-squared: Represents the proportion of the variance in the response
variable explained by the model (higher is better).\
Harris model: R-squared of 0.34, meaning the model explains about 34.6%
of the variance in Harris's polling data.\
Trump model: R-squared of 0.34 as well, indicating similar explanatory
power for Trump's polling data.

@@

Interpretation Summary Predictive Accuracy: The Harris model has
slightly better predictive accuracy than the Trump model, as reflected
by its lower RMSE and MAE values.\
Model Fit: Both models explain roughly 34% of the variance in their
respective datasets. This suggests that other factors not included in
the model may play a significant role in explaining the remaining
variance.\
This summary indicates the models are moderately predictive, with room
for improvement in accuracy and fit, potentially by adding more
predictors or adjusting model specifications.\

@@

# The New York Times/Siena College Polling Methodology {#third-poll}

@@

Siena College Polling Institute is a prominent pollster known for its
comprehensive and methodologically rigorous surveys. It specializes in
political polling and is particularly recognized for its work in
understanding voter sentiment during elections.

Established in 1980 at Siena College in New York's Capital District, the
Siena College Research Institute (SCRI) conducts surveys at regional,
state, and national levels on various topics, including business,
economics, politics, voter behavior, social issues, academics, and
history. The institute carries out both expert and public opinion polls.

Students from Siena and other colleges participate in every survey,
gaining hands-on experience in fields such as political science,
computing, communications, sociology, and psychology. SCRI also hires
interns for special projects involving event planning, in-depth
research, report writing, and analysis. The results of SCRI's surveys
are featured in major regional and national publications, including The
Wall Street Journal and The New York Times, as well as in academic
journals, books, and encyclopedias, both in print and online. Their
findings are regularly highlighted in local and national television and
radio broadcasts. SCRI conducts the Siena New York Poll, a monthly
survey that captures the opinions of registered voters across New York
State on current political issues, along with the New York State Index
of Consumer Sentiment, which offers a quarterly assessment of New
Yorkers' spending intentions. SCRI adheres to the American Association
of Public Opinion Research (AAPOR) Code of Professional Ethics and
Practices. The institute is often commissioned to carry out surveys for
various organizations, businesses, and local and state government
agencies. [@aboutsiena]

@@

We begin by collecting the **eligible voter population** and
**historical voter turnout rates** for different states. In this
simplified example, we will focus on two states: **State A** and **State
B**.

>>>>

---
title: "Forecasting the 2024 U.S. Presidential Election Using a Model-Based Approach"
subtitle: "Harris expects to win on 49%, leads Trump by about 2% in 2024 US election"
author: 
  - Diana Shen
  - Jinyan Wei
  - Jerry Yu
thanks: "Code and data are available at: [https://github.com/DianaShen1224/Forecast-2024-US-election](https://github.com/DianaShen1224/Forecast-2024-US-election)."
date: today
date-format: long
abstract: "This paper predicts the outcome of the 2024 U.S. presidential election using a statistical model based on aggregated polling data for Kamala Harris and Donald Trump. We employ multi-level regression with post-stratification (MRP) using demographic predictors to estimate voter support. The analysis addresses polling biases and proposes an idealized survey methodology with a $100,000 budget. Our results offer insights into voter behavior and suggest improvements for future election forecasting models."
format: pdf
number-sections: true
bibliography: references.bib
editor: 
  markdown: 
    wrap: 72
---

@@

```{r}
#| include: false
#| warning: false
#| message: false
harris_data <- read_parquet(here::here("data", "02-analysis_data", "analysis_data_Harris.parquet"))
trump_data <- read_parquet(here::here("data", "02-analysis_data", "analysis_data_Trump.parquet"))
```

@@

The structure of this paper is organized as follows: @third-data provides
details on the data sources and variables used in our analysis.
@forth-model explains the modeling approach, including the assumptions and
specifications of our linear regression framework. In @fifth-results, we
present our findings, emphasizing the key predictors of Harris's
support. Finally, @sixth-discussion explores the implications of our
results and suggests potential directions for future research. @appendix-a provide external data detail, @appendix-b provide model detail, @appendix-c provide a exploration for pollster methodology, @appendix-d provide a idealized methodology.

@@

In this study, we utilized several R packages to enhance our data
manipulation, modeling, and visualization capabilities. The tidyverse
package offered a comprehensive set of tools for data wrangling and
analysis, improving workflow efficiency [@thereferencecanbewhatever].
The here package aided in managing file paths, allowing for easy access
to our data files [@citehere]. We relied on janitor to perform data
cleaning, as it provides functionalities to identify and rectify quality
issues within the dataset [@citejanitor]. For handling date-related
operations, the lubridate package proved invaluable, simplifying the
manipulation of time variables [@citelubridate]. Lastly, arrow supported
efficient data input and output in a performance-oriented format,
essential for managing larger datasets [@citearrow]. Our coding
practices and file organization were informed by the structure outlined
in @tellingstories.

@@

## Model{#forth-model}

@@

```{r}
#| include: false
#| warning: false
#| message: false
harris_unweighted_model <- readRDS(here::here("models", "model_harris_unweighted.rds"))
harris_weighted_model <- readRDS(here::here("models", "model_harris_weighted.rds"))
trump_unweighted_model <- readRDS(here::here("models", "model_trump_unweighted.rds"))
trump_weighted_model <- readRDS(here::here("models", "model_trump_weighted.rds"))
```

@@

## Predicted Voting Outcomes

```{r}
modelsummary::msummary(list(
  "Harris Unweighted" = harris_unweighted_model, 
  "Harris Weighted" = harris_weighted_model,
  "Trump Unweighted" = trump_unweighted_model,
  "Trump Weighted" = trump_weighted_model))

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
set.seed(233)

new_data_harris <- data.frame(
  end_date = seq(
    min(harris_data$end_date),
    max(harris_data$end_date),
    length.out = 2000
  ),
  state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 2000, replace = TRUE)), 
  pollster = factor(sample(levels(harris_data$pollster), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$pollster)), 
  population = factor(sample(levels(harris_data$population), size = 2000, replace = TRUE), 
                    levels = levels(harris_data$population)),
  sample_size = sample(1:2300, size = 2000, replace = TRUE)
)|>
  mutate(
    national_poll = if_else(state == "National", 1, 0),
    recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),
    combined_weight = recency_weight * sample_size_weight,
    candidate="Harris"
  )

####### Prediction for Trump ######
new_data_trump <- data.frame(
  end_date = seq(
    min(trump_data$end_date),
    max(trump_data$end_date),
    length.out = 2000
  ),
  pollster = factor(sample(levels(trump_data$pollster), size = 2000, replace = TRUE), 
                    levels = levels(trump_data$pollster)),
 population= factor(sample(levels(trump_data$population), size = 2000, replace = TRUE), 
                    levels = levels(trump_data$population)),
   state = factor(sample( c("National","Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin"), size = 2000, replace = TRUE), 
                    levels = levels(trump_data$state)),  
  sample_size =sample(1:2300, size = 2000, replace = TRUE),
  pollster_quality= sample(2:4, size = 2000, replace = TRUE)
)|>
  mutate(
   national_poll = if_else(state == "National", 1, 0),
      recency = as.numeric(difftime(as.Date("2024-11-05"), end_date, units = "days")),
    sample_size_weight = pmin(sample_size / 2300, 1),
    recency_weight = exp(-recency * 0.1),  
    combined_weight = recency_weight * sample_size_weight,
    candidate="Trump"
  )
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| include: false
# Predictions for Harris
new_data_harris <- new_data_harris |>
  mutate(
    predicted_unweighted = predict(harris_unweighted_model, newdata = new_data_harris),
    predicted_weighted = predict(harris_weighted_model, newdata = new_data_harris)
  )

@@

# Filter combined_predictions to include only data from the recent date cutoff onward
combined_predictions<- bind_rows(new_data_harris, new_data_trump)
combined_predictions_filtered<-combined_predictions|>filter(end_date>=recent_date_cutoff)
# Determine the maximum end date for setting the x-axis limit
max_end_date <- as.Date("2024-10-26")
```

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump using Fitted Linear Model"
#| fig-subcap: ["Predicted Support by Unweighetd Model","Predicted Support by Weighetd Model"]
#| layout-ncol: 2
#| label: fig-support-trend
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.7,color="grey")+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using unweighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",caption = 
                   "Linear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))

# Plotting with the filtered data
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.7,color="grey")+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ candidate,ncol = 2) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump in 2024 US Election \n(From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using weighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red"))  +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date))
```
## Predicted Voting Outcomes for Recent Three Months by Respondent Group

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump by respondent group using Fitted Linear Model"
#| fig-subcap: ["Predicted Support by Unweighetd Model","Predicted Support by Weighetd Model"]
#| layout-ncol: 2
#| label: fig-support-respondent
custom_labels <- c("lv" = "Likely Voters", "rv" = "Registered Voters")
### unweighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.7,color="grey")+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ population,ncol = 2, labeller = labeller(population = custom_labels)) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump by respondent group \n in 2024 US Election (From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using unweighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Linear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
### weighted
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.7,color="grey")+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ population,ncol = 2, labeller = labeller(population = custom_labels)) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump by respondent group \n in 2024 US Election (From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using weighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Weights: recency_weight * sample_size_weight \nLinear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 45, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```
## Predicted Voting Outcomes for Recent Three Months by State

```{r}
#| echo: false
#| warning: false
#| message: false
#| eval: true
#| fig-cap: "Predicted Support Trends for Kamala Harris and Donald Trump by State using Fitted Linear Model"
#| fig-subcap: ["Predicted Support by Unweighetd Model","Predicted Support by Weighetd Model"]
#| layout-ncol: 2
#| label: fig-support-respondent
## unweighted
selected_states <- c("Arizona", "Georgia", "Michigan", "Nevada", "North Carolina", "Pennsylvania", "Wisconsin")
combined_data_total<-combined_data_total|>filter(state %in% selected_states)
combined_predictions_filtered<-combined_predictions_filtered|>filter(state %in% selected_states)
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate)) +
  geom_point(alpha = 0.7,color="grey")+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_unweighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump by State \n in 2024 US Election (From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using unweighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       caption = "Linear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
## weighetd
ggplot(combined_data_total,aes(x = end_date,y=pct, colour = candidate,size = combined_weight)) +
  geom_point(alpha = 0.7,color="grey")+ 
  geom_line(data = combined_predictions_filtered,aes(x=end_date,y = predicted_weighted), size = 1)+ # Smoothed trend line
  facet_wrap(~ state,ncol = 4) +
  labs(title = "Predicted Support for Kamala Harris and Donald Trump by State \n in 2024 US Election (From August 1, 2024 to October 27, 2024)",
       subtitle = "predicted using weighted multiple linear regression model ",
       x = "End Date of the Poll",  # Change x-axis label to End Date
       y = "Predicted Support (%)",
       color = "Candidate",
       size="Weight",
       caption = "Linear Regression Model: Candidate Support Percentage =national_poll + pollster + population") +
  scale_color_manual(values = c("Harris" = "blue", "Trump" = "red")) +
  theme_classic() +
  theme(legend.position = "top",
        axis.text.x = element_text(angle = 90, hjust = 1),strip.background = element_rect(colour = "black", fill = "white")) +
  scale_x_date(date_labels = "%b %d", date_breaks = "1 week", limits = c(recent_date_cutoff, max_end_date)) 
```

```{r}
# Load necessary package
library(knitr)
library(kableExtra)

# Example: Assume `predicted_results` is your data frame
# Ensure column names match this format
# Example structure:
# predicted_results <- data.frame(
#   State = c("Arizona", "California", ...),
#   Trump_Predicted_Percentage = c(48.94, 33.80, ...),
#   Electoral_Votes = c(11, 55, ...),
#   Winner = c("Harris", "Harris", ...)
# )

# Rename columns to match the desired table format
colnames(predicted_results) <- c("State", "Harris Predicted %", "Vote Share", "Winner")

# Generate table
predicted_results %>%
  kbl(caption = "Prediction Vote share for Trump and Harris", format = "html") %>%
  kable_styling(full_width = FALSE, position = "center") %>%
  column_spec(2, width = "5em") %>%
  column_spec(3, width = "5em") %>%
  column_spec(4, width = "5em") %>%
  add_header_above(c(" " = 1, "Trump Predicted %" = 1, "Electoral Votes" = 1, "Winner" = 1))

@@

# Additional data details {#appendix-a}

@@

# Model details {#appendix-b}

@@

Harris model: RMSE of 3.42, indicating an average prediction error of
around 3.44 percentage points.\
Trump model: RMSE of 4.30, showing a slightly higher prediction error on
average.\

R-squared: Represents the proportion of the variance in the response
variable explained by the model (higher is better).\
Harris model: R-squared of 0.33, meaning the model explains about 34.6%
of the variance in Harris's polling data.\
Trump model: R-squared of 0.27 as well, indicating similar explanatory
power for Trump's polling data.

@@

Interpretation Summary Predictive Accuracy: The Harris model has
slightly better predictive accuracy than the Trump model, as reflected
by its lower RMSE and MAE values.\
Model Fit: Harris' model explain roughly 33% of the variance in their
respective datasets. Trump's models explain roughly 33% of the variance in their
respective datasets.This suggests that other factors not included in
the model may play a significant role in explaining the remaining
variance.\
This summary indicates the models are moderately predictive, with room
for improvement in accuracy and fit, potentially by adding more
predictors or adjusting model specifications.\

@@

# The New York Times/Siena College Polling Methodology {#appendix-c}

@@

Siena College Polling Institute is a prominent pollster known for its
comprehensive and methodologically rigorous surveys. It specializes in
political polling and is particularly recognized for its work in
understanding voter sentiment during elections. Established in 1980 at
Siena College in New York's Capital District, the institute carries out
both expert and public opinion polls. [@aboutsiena].

@@

We begin by gathering data on the **eligible voter population** and
**historical ballots cast** for different states. For simplicity, we'll
focus on two states: **State A** and **State B**.

| State   | Eligible Voters | Ballots Cast |
|---------|-----------------|--------------|
| State A | 1,000,000       | 500,000      |
| State B | 1,000,000       | 700,000      |

### Step 2: Calculate Total Ballots Cast
